\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}

### 2022-05-13 sections/discussion.tex
Change: Updates based on Reviewer 2 comments

  The current analyses make use of one fecal microbiome transplant dataset with healthy and autistic samples, three mock community datasets and several datasets generated by two synthetic interaction methods.
  The analyses presented in this study are specific to the datasets utilized, and due to the lack of a gold standard reference dataset for microbial interaction data, it is not possible to generalize these results to datasets with widely different microbiome profiles.
  Incorporating datasets from a broad spectrum of biomes with varying microbial distributions into \ac{micone} will likely increase the robustness and generalizability of the results from these analyses.
  Additionally, the utilization of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a} as additional sources for benchmarking
  Furthermore, expanding the scope of the \ac{micone} pipeline to also include metagenomics sequencing data will enable an even broader access to a diverse range of datasets as well as domains of life that can be incorporated into the analysis.

>>>>

  The current analyses make use of one fecal microbiome transplant dataset with healthy and autistic samples, three mock community datasets and several datasets generated by two synthetic interaction methods.
  The analyses presented in this study are specific to the datasets utilized, and due to the lack of a gold standard reference dataset for microbial interaction data, it is not possible to generalize these results to datasets with widely different microbiome profiles.
  Incorporating datasets from a broad spectrum of biomes with varying microbial distributions into \ac{micone} will likely increase the robustness and generalizability of the results from these analyses.
  The network analyses in this study are primarily at the Genus level, wherein the lowest resolution of a node is a Genus and if an entity cannot be resolved to the Genus level and next lowest taxonomic level is used (e.g. Family).
  As a consequence, two entities belonging to the same lineage where one entity is resolved to the Genus level and another is resolved to the Family level are treated as two different nodes in the network.
  Therefore, the development of a metric of overlap to compare nodes with shared lineages within and across networks could provide more biologically and phylogentically relevant comparisons.
  Further analyses could also include networks at the Species or other taxonomic levels.

  Additionally, the utilization of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a} as additional sources for benchmarking
  Furthermore, expanding the scope of the \ac{micone} pipeline to also include metagenomics sequencing data will enable an even broader access to a diverse range of datasets as well as domains of life that can be incorporated into the analysis.

### 2022-05-13 sections/results.tex
Change: Updates based on Reviewer 2 comments

  These comparisons only elucidate the pairwise similarity or dissimilarity of a pair of methods.
  In order to determine the tool that most accurately recapitulates the reference sequences in the samples, we used the information about the known 16S sequences from the mock datasets (mock4, mock12, and mock16).
  In particular, we used the pipeline to process mock community datasets using each of the possible methods included for this step.
  Subsequently, we compared the predicted representative sequences with the expected representative sequences and their distribution.
  The results (Figure~\ref{fig:figure3}C and D) show that, for the mock datasets, the predicted sequence distributions are substantially different from the expected sequence distribution.
  This can be seen in the large values for both unweighted and weighted UniFrac distance values when compared to Figure~\ref{fig:figure3}A and B.
  This result is more exaggerated in the case of the unweighted UniFrac metric, where some processed mock datasets show a very high deviation from the expected sequences.
  These high deviations are primarily in one or two of the three datasets that were analyzed and show that the datasets themselves play a big role in the performance of these methods.
  Open-reference and de novo clustering methods perform the best under the weighted UniFrac metric and the worst (marginally) under the unweighted UniFrac metric.
  This result can be attributed to the large number of low abundance representative sequences that are generated by these methods.
  Deblur performs poorly under weighted Unifrac and although its performance on the mock4 dataset is the best under unweighted UniFrac, its performance on the other datasets is average.
  The Deblur method returns a very small number of representative sequences (2388) and this could account for the reason for the high dissimilarity with the other methods as well as irregular performance on the mock data.
  There is no method that clearly outperforms the rest in all datasets.
  Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies~\cite{Nearing2018}, \ac{dada2} and Deblur appear to be the most reliable.
  Overall, since \ac{dada2} as compared to Deblur, displays better performance on all the mock datasets and under the weighted UniFrac metric, we have decided to recommend this tool for the DC step of the pipeline.

>>>>

  These comparisons only elucidate the pairwise similarity or dissimilarity of a pair of methods.
  In order to determine the tool that most accurately recapitulates the reference sequences in the samples, we used the information about the known 16S sequences from the mock datasets (mock4, mock12, and mock16).
  In particular, we used the pipeline to process mock community datasets using each of the possible methods included for this step.
  Subsequently, we compared the predicted representative sequences with the expected representative sequences and their distribution.
  The results (Figure~\ref{fig:figure3}C and D) show that, for the mock datasets, the predicted sequence distributions are substantially different from the expected sequence distribution.
  This can be seen in the large values for both unweighted and weighted UniFrac distance values when compared to Figure~\ref{fig:figure3}A and B.
  This result is more exaggerated in the case of the unweighted UniFrac metric, where some processed mock datasets show a very high deviation from the expected sequences.
  These high deviations are primarily in one or two of the three datasets that were analyzed and show that the datasets themselves play a big role in the performance of these methods.
  Open-reference and de novo clustering methods perform the best under the weighted UniFrac metric and the worst (marginally) under the unweighted UniFrac metric.
  This result can be attributed to the large number of low abundance representative sequences that are generated by these methods.
  Deblur performs poorly under weighted Unifrac and although its performance on the mock4 dataset is the best under unweighted UniFrac, its performance on the other datasets is average.
  The Deblur method returns a very small number of representative sequences (2388) and this could account for the reason for the high dissimilarity with the other methods as well as irregular performance on the mock data.
  There is no method that clearly outperforms the rest in all datasets.
  Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies~\cite{Nearing2018}, \ac{dada2} and Deblur appear to be the most reliable.
  This is because the open-reference and de novo clustering methods return a much larger number of \ac{otu}s compared to the other pipelines and this would affect the accuracy of the network inference step if filtering is not performed.
  Overall, since \ac{dada2} as compared to Deblur, displays better performance on all the mock datasets and under the weighted UniFrac metric, we have decided to recommend this tool for the DC step of the pipeline.
  However, if comparison across studies that have sequenced different 16S regions is required, closed-reference and open-reference might be a better option.

### 2022-05-13 sections/tables_figures.tex
Change: Updates based on Reviewer 2 comments

  % TODO: Add equations for the linear model to the left of the pie chart
  \begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{figure2.pdf}
  \end{figure}
  \begin{figure}[H]
    \centering
      \caption{
      \textbf{The choice of reference database contributes to the most variance in the networks}.
      \textbf{(A)} The percentage of variance in the networks contributed by the DC, CC (chimera checking), TA and OP steps of the pipeline (right) and the linear model used to calculate the variance (left), see the Methods section for details.
      The data used for the analysis were the healthy stool samples from the fecal microbiome transplant dataset~\ref{Kang2017}.
      The taxonomy database contributes most to the variance between the networks (65.4\%) followed by the filtering of the counts matrix (26.8\%) in the OP step.
      The variation due to the DC and CC steps are comparatively not as significant.
      \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.
      Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
      The color of the points corresponds to the reference taxonomy database used to reconstruct those networks (left), and the network inference algorithm used to infer the networks (right).
      The points on the PCA plot can be grouped based on the TA step, but at the NI step only some methods create high variability in the inferred networks.
    }
    \label{fig:figure2}
  \end{figure}
  \FloatBarrier
  \newpage

  \begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{figure3.pdf}
  \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ in the sequences that are lower in abundance.}
      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods (with the exception of Deblur (DB)).
      \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences that are lower in abundance.
      The number of \ac{otu}s or \ac{esv}s generated by the respective methods are provided in the parenthesis next to the names.
      \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the predicted sequence profile in the mock datasets.
      The distributions of the average weighted UniFrac distance show that de novo (DN) and open reference (OR) were the best performing methods in most of the datasets, while they are worst performing methods under the unweighted UniFrac metric.
      The good performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to propose it as the default method for the DC step.
    }
    \label{fig:figure3}
  \end{figure}
  \FloatBarrier
  \newpage

  \begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=1.2\textwidth]{figure4.pdf}
  \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Taxonomic reference databases vary widely in terms of their taxonomy assignments beyond the Order level.}
      \textbf{(A)} The assignments of the top 50 representative sequences to their respective taxonomies using the three different reference databases.
      This result clearly describes how the same sequences are assigned to different Genus under the different databases.
      A significant portion of the representative sequences are assigned to an ``unknown'' Genus in two of three databases (\ac{gg} and \ac{ncbi}).
      \textbf{(B)} The number of representative sequences assigned to the same taxonomic label when using different reference databases, shown for the top 100 sequences.
      The number of mismatches are fewer at higher taxonomic levels, but, even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
      However, we still recommend the \ac{gg} database and the Naive Bayes classifier as the defaults for the \ac{micone} pipeline.
      Due to the popularity of the \ac{gg} database, comparison against existing networks in literature will be easier since many published datasets make use of this database in their analyses.
    }
    \label{fig:figure4}
  \end{figure}
  \FloatBarrier
  \newpage

  \begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{figure5.pdf}
  \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
      \textbf{(A)} The nine different networks (excluding \ac{mldm}) generated by the different network inference methods are very dissimilar.
      The green links are positive associations and the orange links represent negative associations.
      A threshold of 0.3 was set for the correlation-based methods (\ac{sparcc}, Propr, Spearman and Pearson) and a threshold of 0.01 was set for the other direct association methods.
      The correlation-based methods in general produce graph with higher edge-densities.
      \textbf{(B)} The node overlap Upset plot indicates that all the networks have a large proportion of common nodes involved in connections (33 out of 68).
      Whereas, \textbf{(C)} the edge overlap Upset plot shows that a very small fraction of these connections are actually shared (8 out of 202).
    }
    \label{fig:figure5}
  \end{figure}
  \FloatBarrier
  \newpage

@@

  % TODO: Verify the validity of Figure 7A and 7B again
  % Are they both using consensus?
  \begin{figure}[H]
    \centering
    \includegraphics[width=1.0\linewidth]{figure7.pdf}
  \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{The consensus algorithm drastically reduces variability in the networks due to upstream data processing methods and increases robustness of inferred networks}
      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, CC=remove bimera, TA=\ac{gg}, OP=Filter(on), NI=scaled-sum consensus) is compared with networks generated when one of the steps use a different tool.
      The common connections (common with the default network) are in black, connections unique to the network are colored purple and connections in the default network, but not present in the current network are gray.
      Changing the NI, TA and OP steps leads to the creation of the most number of new nodes and edges.
      \textbf{(B)} The boxplot showing the L1 distance between the default network and the networks generated by changing one step of the default pipeline.
      The high value of L1 distances for the network inference step as compared to the other steps demonstrates how the consensus method is able to reduce the variability in the networks in the other steps of the pipeline.
      This also shows that even though many new nodes were added in (A) when the TA and OP steps were changed, the L1 distances remain fairly small.
      \textbf{(C)} The networks generated for the control (left) and autistic (right) samples in the fecal microbiome transplant dataset.
      These networks were generated using the default tools and parameters recommended by the \ac{micone} pipeline.
    }
    \label{fig:figure7}
  \end{figure}

>>>>

  % TODO: Add equations for the linear model to the left of the pie chart
  \begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{figure2.pdf}
  \end{figure}
  \begin{figure}[H]
    \centering
      \caption{
      \textbf{The choice of reference database contributes to the most variance in the networks}.
      \textbf{(A)} The percentage of variance in the networks contributed by the DC, CC (chimera checking), TA and OP steps of the pipeline (right) and the linear model used to calculate the variance (left), see the Methods section for details.
      The data used for the analysis were the healthy stool samples from the fecal microbiome transplant dataset~\cite{Kang2017}.
      The taxonomy database contributes most to the variance between the networks (65.4\%) followed by the filtering of the counts matrix (26.8\%) in the OP step.
      The variation due to the DC and CC steps are comparatively not as significant.
      \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.
      Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
      The color of the points corresponds to the reference taxonomy database used to reconstruct those networks (left), and the network inference algorithm used to infer the networks (right).
      The points on the PCA plot can be grouped based on the TA step, but at the NI step only some methods create high variability in the inferred networks.
    }
    \label{fig:figure2}
  \end{figure}
  \FloatBarrier
  \newpage

  \begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{figure3.pdf}
  \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ in the sequences that are lower in abundance.}
      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods (with the exception of Deblur (DB)).
      \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences that are lower in abundance.
      The number of \ac{otu}s or \ac{esv}s generated by the respective methods are provided in the parenthesis next to the names.
      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the fecal microbiome transplant dataset.
      \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the predicted sequence profile in the mock datasets.
      The distributions of the average weighted UniFrac distance show that de novo (DN) and open reference (OR) were the best performing methods in most of the datasets, while they are worst performing methods under the unweighted UniFrac metric.
      The good performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to propose it as the default method for the DC step.
      The data used for the analysis in \textbf{(C, D)} were the mock datasets from mockrobiota~\cite{Bokulich2016}.
    }
    \label{fig:figure3}
  \end{figure}
  \FloatBarrier
  \newpage

  \begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=1.2\textwidth]{figure4.pdf}
  \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Taxonomic reference databases vary widely in terms of their taxonomy assignments beyond the Order level.}
      \textbf{(A)} The assignments of the top 50 representative sequences to their respective taxonomies using the three different reference databases.
      This result clearly describes how the same sequences are assigned to different Genus under the different databases.
      A significant portion of the representative sequences are assigned to an ``unknown'' Genus in two of three databases (\ac{gg} and \ac{ncbi}).
      \textbf{(B)} The number of representative sequences assigned to the same taxonomic label when using different reference databases, shown for the top 100 sequences.
      The number of mismatches are fewer at higher taxonomic levels, but, even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the fecal microbiome transplant dataset.
      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
      However, we still recommend the \ac{gg} database and the Naive Bayes classifier as the defaults for the \ac{micone} pipeline.
      Due to the popularity of the \ac{gg} database, comparison against existing networks in literature will be easier since many published datasets make use of this database in their analyses.
      The datasets used for the analysis in \textbf{(C)} were the mock datasets from mockrobiota.
    }
    \label{fig:figure4}
  \end{figure}
  \FloatBarrier
  \newpage

  \begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{figure5.pdf}
  \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
      \textbf{(A)} The nine different networks (excluding \ac{mldm}) generated by the different network inference methods are very dissimilar.
      The green links are positive associations and the orange links represent negative associations.
      A threshold of 0.3 was set for the correlation-based methods (\ac{sparcc}, Propr, Spearman and Pearson) and a threshold of 0.01 was set for the other direct association methods.
      The correlation-based methods in general produce graph with higher edge-densities.
      \textbf{(B)} The node overlap Upset plot indicates that all the networks have a large proportion of common nodes involved in connections (33 out of 68).
      Whereas, \textbf{(C)} the edge overlap Upset plot shows that a very small fraction of these connections are actually shared (8 out of 202).
      The data used in this analysis were the healthy stool samples from the fecal microbiome transplant dataset.
    }
    \label{fig:figure5}
  \end{figure}
  \FloatBarrier
  \newpage

@@

  % TODO: Verify the validity of Figure 7A and 7B again
  % Are they both using consensus?
  \begin{figure}[H]
    \centering
    \includegraphics[width=1.0\linewidth]{figure7.pdf}
  \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{The consensus algorithm drastically reduces variability in the networks due to upstream data processing methods and increases robustness of inferred networks}
      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, CC=remove bimera, TA=\ac{gg}, OP=Filter(on), NI=scaled-sum consensus) is compared with networks generated when one of the steps use a different tool.
      The common connections (common with the default network) are in black, connections unique to the network are colored purple and connections in the default network, but not present in the current network are gray.
      Changing the NI, TA and OP steps leads to the creation of the most number of new nodes and edges.
      \textbf{(B)} The boxplot showing the L1 distance between the default network and the networks generated by changing one step of the default pipeline.
      The high value of L1 distances for the network inference step as compared to the other steps demonstrates how the consensus method is able to reduce the variability in the networks in the other steps of the pipeline.
      This also shows that even though many new nodes were added in (A) when the TA and OP steps were changed, the L1 distances remain fairly small.
      \textbf{(C)} The networks generated for the control (left) and autistic (right) samples in the fecal microbiome transplant dataset.
      These networks were generated using the default tools and parameters recommended by the \ac{micone} pipeline.
      The data used in this analysis were the healthy stool samples from the fecal microbiome transplant dataset.
    }
    \label{fig:figure7}
  \end{figure}
